Measuring Perceptions and Preferences for Meritocracy

### 2020-09-15 docs/paper.Rmd
Revisión y conclusiones

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequality have raised growing concerns and crises across societies. One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on effort and talent. Despite its wide use, empirical research on meritocracy is something relatively novel. A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items. Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy. Based on this critical analysis, we propose an item's battery called *Perceptions and preferences for meritocracy scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141). The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::pdf_document2:
    template: null
    toc: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962]. Esta concepción tradicional de mérito pone en un lugar secundario la posible interferencia de factores estructurales o no meritocráticos como la herencia, los contactos personales, o la suerte [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994]. Una serie de estudios han planteado críticas a la realización de este estándar moral de distribución, argumentando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019]. Por otro lado, desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].

@@

Para poder dar cuenta de los niveles de creencia en la meritocracia los estudios a la fecha generalmente han utilizado algunos indicadores de encuestas ya existentes, y en el menor de los casos se han creado instrumentos ad-hoc. Sin embargo, y como mostraremos más adelante, las formas de medición de meritocracia varían extremadamente entre estudios. Muchas veces fenómenos similares se asocian a indicadores distintos, y también ocurre que fenómenos distintos son medidos con indicadores similares, todo lo cual hace dificulta la comparabilidad entre estudios y el poder avanzar en la comprensión y estudio de la meritocracia.

Basados en el análisis crítico de las formas de medición de meritocracia a la fecha, el presente artículo propone un instrumento para medir y relacionar dos aspectos claves en el estudio de la meritocracia: percepciones y preferencias. Además, como un segundo eje de análisis considera la generación de indicadores respecto de aspectos meritocráticos y anti-meritocráticos, demostrando que no son los dos polos de un mismo continuo como muchos estudios anteriores parecen sugerir. La propuesta de medición además está orientada a generar un instrumento lo más breve posible de manera que pueda ser utilizado en encuestas de opinión pública y así ser asociado a otros fenómenos sociales.

@@

Several approaches to the empirical study of meritocracy based on public opinion surveys make reference to the concept of _beliefs_, but behind this concept there are usually different meanings and operationalizations. To illustrate this point in the following we will start with the proposal from a recent paper by @mijs_paradox_2019, which we will take as a reference to discuss previous studies.

The meritocratic beliefs' definition of Mijs is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9]. In the operationalization, this is associated with the following indicator: "how important you think it is for getting ahead in life: (a) hard work", scored in a 1 to 5 likert scale. There are several assumptions behind this decision that are worth discussing:

@@

The item used by Mijs is part of an items' battery present in several international surveys, usually called "reasons to get ahead". This battery presents a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender. Therefore, for Mijs other aspects as education, that could be associated to talent, are not meritocratic. As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation, ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is conditioned by a nonmeritocratic factor: who your parents happen to be" (p.5).

In Mijs' proposal we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization). The problem of whether talent is or not meritocratic is something that actually should (and can) be tested, it cannot be simply ruled out particularly in empirical research. This conceptual and measurement issue it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].

@@

The "resons to get ahead" battery refers to "how important you think it is", considered by Mijs as a belief in meritocracy for the item regarding effort. Nevertheless, another version of this same battery in the survey asks about "how important you think it _should_ be".  Which one of both is a "belief"? According to @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e. thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e. thoughts about what should be)"(p.359). Therefore, the referred paper would be actually using the term beliefs referring to perceptions. The same occurs in @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use different terms as attitudes (@kunovich_systems_2007). The first attempt to shed light on this confusion was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “... people are rewarded for their skills?”.

Is the belief in meritocracy a perception or a desire/preference? In order to expand the analitical conceptual framework, we believe that both dimensions should be included in the analysis, as proposed by @duru-bellat_whos_2012. This opens up possibilities of analyzing whether perceptions and preferences are actually the same (i.e. correlation close to 1) or they are different aspects of the same phenomenon. As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p. 435). In other words, normative beliefs should be interpreted in the context of perception: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy. In order to avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences and meritocratic perceptions, as they better reflect the two dimension under scrutiny.

@@

@mijs_paradox_2019 makes reference to some non-meritocratic aspects as talent, which is ruled out of the operationalization of meritocracy. A different approach was followed by @kunovich_systems_2007, who decide to include some non-meritocratic elements. Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), he decided that reasons as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e. reverse coded). A similar approach was taken by @newman_false_2015, reverse-coding non-meritocratic items, similarly to what occurs with the "Preference for the Merit Principle Scale" [@davey_preference_1999].

The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the same "get ahead" perceptions' battery items. They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections and non-mertitocratic. Nevertheless, despite making this distinction the author ends up subtracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did. Still, taking into account this research perspective, we suggest that non-meritocratic aspects should be part of a meritocratic measurement but taken independently and not adding or subtracting from meritocratic ones unless it is empirically proved that they belong to the same conceptual dimension.

@@

Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989], using single indicators and/or simple average indexes for measuring meritocracy. Such strategy assumes that the construct is measured perfectly by the indicators chosen, going as far as proposing that "... In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654]. Some advances were done by @reynolds_perceptions_2014 by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily they end up in a sum index despite proving multidimensionality.

@@

Based on the previous limitations and confusions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:

- Multidimensional, incorporating previous distinctions between preferences and perceptions as well as between meritocratic and non-meritocratic aspects.

- Multiple indicators for each dimension, in order to account for measurement error in a confirmatory factor analysis context.

- Based on previous indicators as far as possible, for the sake of comparability between studies

- Brief, as to be used in a regular survey. In this point it differs for instance from the proposal of the "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items just for one dimension (besides the problem of reverse-coding non-meritocratic items).

@@

The columns Perceptions and Preferences represent the distinction between this two concepts, usually confused under the label "beliefs". Perceptions refers to the extent to which people see that meritocracy functions or applies in their society, which in terms of measurement relates to items such as "I think hard work is important to get ahead in society", whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g. whether hard work should be related to payment). The rows in the table of Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014]. This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny. These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.

@@

The research hypotheses behind this conceptualization are the following:

@@

The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into the 4 dimensions referred above: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic). In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which are mostly used for operationalizing meritocracy in previous empirical studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014]. The eight items ordered according to dimensions are presented in Table \@ref(tab:table-indicadores). These 8 likert-type items have 5 response alternatives ranging from "Completely disagree"(1) to "Completely agree" (5).

@@

With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different order versions as explained in Figure \@ref(fig:appmod) . The scale was presented to the first group (_n = 712_) in the order that appears in Table 2. For the second group (_n = 717_), the order of the items was organized according to the topics of the items, e.g. for the topic of hard work the item about perception was followed by the item about preference, and te same for the rest of the topics. Finally, for the third group (_n = 712_) the items were completely randomized.

@@

Considerando la ventaja del orden, respecto aleatorio despeja la posibilidad del efecto framing, es decir, de que el resultado del modelo se deba al orden de las preguntas, se ha decidido seguir utilizando el orden 3.

uso conjunto percepciones-preferencias

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequality have raised growing concerns and crises across societies. One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on criteria such as effort and talent. Despite its wide use, empirical research on meritocracy is something relatively novel. A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items. Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy. Based on this critical analysis, we propose an item's battery called *Perceptions and Preferences for Meritocracy Scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141). The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::pdf_document2:
    template: null
    toc: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962]. Esta concepción tradicional de mérito sitúa en un lugar secundario la posible interferencia de factores estructurales o no meritocráticos como la herencia, los contactos personales, o la suerte [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994]. Desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019]. Tales investigaciones han planteado críticas a la realización de este estándar moral de distribución, argumentando que sería una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019, @khan_Privilege_2013].

@@

Para poder dar cuenta de los niveles de creencia en la meritocracia los estudios a la fecha generalmente han utilizado algunos indicadores de encuestas ya existentes, y en menor medida se han creado instrumentos ad-hoc. Sin embargo, y como mostraremos más adelante, las formas de medición de meritocracia varían extremadamente entre estudios. Muchas veces fenómenos similares se asocian a indicadores distintos, y también ocurre que fenómenos distintos son medidos con indicadores similares, todo lo cual dificulta la comparabilidad entre investigaciones y avanzar en la comprensión y estudio de la meritocracia.

Basados en el análisis crítico de las formas de medición de meritocracia a la fecha, el presente artículo propone un instrumento para medir y relacionar dos aspectos claves en el estudio de la meritocracia: percepciones y preferencias. Además, como un segundo eje de análisis considera la generación de indicadores respecto de aspectos meritocráticos y anti-meritocráticos, demostrando que no son los dos polos de un mismo continuo como muchos estudios anteriores parecen sugerir. La propuesta de medición además está orientada a generar un instrumento lo más breve posible de manera que pueda ser utilizado en encuestas de opinión pública y así ser estudiado en relación a otros fenómenos sociales.

@@

Several approaches to the empirical study of meritocracy based on public opinion surveys make reference to the concept of _beliefs_, but behind this concept there are usually different meanings and operationalizations. To illustrate this point in the following we will start with the proposal from a recent paper by @mijs_paradox_2019, which we will take as a reference to discuss previous studies on meritocracy.

The meritocratic beliefs' definition of Mijs is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9]. In the operationalization, this is associated with the following indicator: "how important you think it is for getting ahead in life: (a) hard work", scored in a 1 to 5 likert scale. There are several assumptions behind this decision that are worth discussing and that are related to previous assumptions in the study of meritocracy.

@@

The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead". This battery presents a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender. Therefore, for Mijs other aspects such as education, that could be associated to talent, are not meritocratic. As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation, ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is conditioned by a nonmeritocratic factor: who your parents happen to be" (p.5).

In Mijs' proposal we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization). The problem of whether talent is or not meritocratic is certainly an interesting point to discuss, but for those working based on empirical data it is something that should be empirically tested. Is effort the only dimension behind the concept of meritocracy, or is it a multidimensional concept that includes other elements as talent (as in its original conceptualization)? This conceptual and measurement unidimensional assumption it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].

@@

The "reasons to get ahead" battery refers to "how important you think it is", considered by @mijs_paradox_2019 as a belief in meritocracy for the item regarding effort. Nevertheless, another version of this same battery used in several surveys is "how important you think it _should_ be".  Therefore, the question raised here is: Which one of both is a "belief"? According to @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e. thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e. thoughts about what should be)"(p.359). Therefore, the referred paper would be using the term beliefs (what should be) referring actually to perceptions (what is). The same occurs in studies as the one by @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use general terms as attitudes (@kunovich_systems_2007). The first attempt to shed light on this confusion was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “... people are rewarded for their skills?”.

Is the belief in meritocracy a perception or a desire/preference? In order to expand the analytical conceptual framework, we believe that both dimensions should be included in the analysis, as proposed by @duru-bellat_whos_2012. This opens up possibilities of analyzing whether perceptions and preferences are actually related  (i.e. correlation close to 1) or whether they are different aspects of the same phenomenon. As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p. 435). In other words, normative beliefs should be interpreted taken perception into account: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy. In order to avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences and meritocratic perceptions, as they better reflect the two dimension under scrutiny.

@@

@mijs_paradox_2019 makes reference to some non-meritocratic aspects as talent, which is ruled out of the operationalization of meritocracy. A different approach was followed by @kunovich_systems_2007, who decide to include some non-meritocratic elements. Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), they decided that reasons as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e. reverse coded). A similar approach was taken by @newman_false_2015, reverse-coding non-meritocratic items, the same principle applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].

The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the  "get ahead" perceptions' battery items mentioned above. They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections and non-mertitocratic. Nevertheless, despite making this distinction the author ends up subtracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did. Still, taking into account this research perspective, we suggest that non-meritocratic aspects should be part of a meritocratic measurement but taken independently and not adding or subtracting from meritocratic ones unless it is empirically proved that they belong to the same conceptual dimension.

@@

Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989], using single indicators and/or simple average indexes for measuring meritocracy. Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as proposing that "... In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654]. Some advances were done by @reynolds_perceptions_2014 by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they end up in a sum index despite proving a multidimensional latent structure.

@@

Based on the previous limitations and assumptions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:

- _Multidimensional_, incorporating previous distinctions between preferences and perceptions as well as between meritocratic and non-meritocratic aspects.

- Multiple indicators for each dimension, in order to _account for measurement error_ in a confirmatory factor analysis context.

- Based on _previous indicators_ as far as possible, for the sake of comparability between studies

- _Brief_, as to be used in regular public opinion surveys. In this point it differs for instance from the proposal of the "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items just for one dimension (besides the problem of reverse-coding non-meritocratic items).

@@

The columns Perceptions and Preferences represent the distinction between this two concepts, usually confused under the label "beliefs". Perceptions refers to the extent to which people observe that meritocracy functions or applies in their society, which in terms of measurement relates to items such as "I think hard work is important to get ahead in society", whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g. whether hard work should be related to payment). The rows in the table of Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014]. This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny. These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.

@@

The research hypotheses behind this conceptualization and measurement model are the following:

@@

The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into the 4 dimensions referred above: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic). In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which are mostly used for operationalizing meritocracy [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014]. The eight items ordered according to dimensions are presented in Table \@ref(tab:table-indicadores). These 8 likert-type items have 5 response alternatives ranging from "Completely disagree"(1) to "Completely agree" (5).

@@

With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different order versions as explained in Figure \@ref(fig:appmod). The scale was presented to the first group (_n = 712_) in the order that appears in Table 2. For the second group (_n = 717_), the order of the items was organized according to the topics of the items, e.g. for the topic of hard work the item about perception was followed by the item about preference, and te same for the rest of the topics. Finally, for the third group (_n = 712_) the items were completely randomized.

@@

Studies that attempt to characterize and compare societies regarding support for meritocratic beliefs have used different approaches. As most studies use secondary survey data, they tend to grasp in the available indicators assuming that they represent an underlying construct. Nevertheless, a review of these studies reveal several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy. As the existence of heterogeneous approaches certainly have consequences for advancing in the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy built upon previous research, which is then tested again new survey data.

The results of the analysis indicate that is possible to distinguish between different dimensions in the study of meritocracy based on the proposed 8-items scale Perceptions And Preferences for Meritocracy. On the one hand, perceptions and preferences proved to be two related but different dimensions, usually confounded in previous research under the label of "beliefs". On the other hand, meritocratic and non-meritocratic dimensions are not poles of a same continuum as previosuly assumed.

The four-dimensional framework and structure of the Perceptions and Preferences for Meritocracy scale opens several avenues for future research. For instance, distinguishing perceptions from preferences will allow to evaluate the extent to which different societies are accustomed or satisfied with the performance of meritocracy, in terms of difference of what is perceived from what is preferred. On the other hand, given that non-meritocratic elements are not necessarily related to the meritocratic ones, this open the door to assess the legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies. The consideration of structural factors, both individual and societal, will allow to advance in hypotheses of meritocratic legitimacy from a more comprehensive view than the used today based only in meritocratic perceptions. Furthermore, the impacts of different configurations of the four dimensional framework on practices and behaviors such as corruption, civic involvement and political views is an area that requires additional research particularly in times of economic crisis and growing inequality, and to which the proposed instrument aims at contributing.
